---
title: "HMK 08"
format: gfm
---

# Joins

To prepare for Question 1, load packages and create flights2:  
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)

flights2 <- flights |>
  select(year, time_hour, origin, dest, tailnum, carrier)
```

1. Imagine you’ve found the top 10 most popular destinations using this code:
```{r}

top_dest <- flights2 |>
  count(dest, sort = TRUE) |>
  head(10)

```
How can you find all flights to those destinations?

To find all the flights to those destinations, I can use the left_join function to combine the top_dest data set with the larger flights2 data set and join only according to dest. I added count(dest, sort = TRUE) to make sure my output was in the same order as the top_dest output. 

```{r warning=FALSE}

all_flights <- top_dest |>
  left_join(flights2, by = "dest") |>
  count(dest, sort = TRUE) 

```


I used nrow() to ensure my all_flights had the correct number of rows. 

```{r}
nrow(all_flights)
```


To be extra certain that they matched, I used the count() function to compare the flights listed in all_flights vs top_dest.
```{r}
all_flights |> count(dest)
```


# Functions

2. Write a function to 'rescale' a numeric vector by subtracting the mean of the vector from each element and then dividing each element by the standard deviation. 

```{r}
r <- c(1:25)

rescale <- function(x) {
  x - mean(x) / sd(x)
}

r2 <- rescale(r)

```
I created a vector of the valued 1-25, which are assigned as an object (r). I chose a name for the function, which i called "rescale," then I listed the arguments to the function inside function. The argument was to subtract the mean of x from x, then divide each value by the standard deviation of x. The argument code is within the body of the function, represented by a { that immediately follows function(). To rescale r I created the object r2 and used rescale() to apply the arguments to that object. 
