---
title: "Sept25_ClassAssignment"
format: html
editor: visual
---

```{r, message=FALSE, warning=FALSE}

fen_counts <- read.csv("fake_eagle_nest_counts.csv")

fen_counts

```

------------------------------------------------------------------------

I would not describe this data as tidy because the each variable is not listed in a column and each observation is not listed in a row. A more useful way to view the data would be having a column for year with all the years listed below, a column for the region with the region specified as the value, and a column for the count of eagles nests present.

So, we will "pivot" the data frame to make it "tidy".

## Pivoted Data Frame

```{r, message=FALSE, warning=FALSE}
library(tidyverse)

fen_counts |>
  pivot_longer(
    cols = !c(year), 
    names_to = "Region", 
    values_to = "Count"
  )

```

## Scatter Plot 

Now the data is tidy and ready to plot. 

```{r}
library(ggplot2)

fen_plot <- fen_counts |>
  pivot_longer(
    cols = !c(year), 
    names_to = "Region", 
    values_to = "Count"
  )

  ggplot(fen_plot, aes(x = Count, y = year, color = Region)) +
  geom_point() + 
    geom_smooth(formula = y~x, method = "lm")
   

```

